---
title: "Data Wrangling (1)"
author: "Nisaar Hussain"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Organization
Always open the R project icon, do not open just the file
R will then be able to recognize it as a root folder
You can use getwd() function to find out the working directory

## Opening Remarks
1. Always load the library first
2. Then load the data files

## Objectives of this Lecture

This lecture introduces data wrangling with R. Using V-Dem data as an example, we will learn how to use the wrangle data with a set of [`tidyverse`](https://www.tidyverse.org/) functionality. Specifically, we will focus on functions...

1.  to import and export data: `read_csv` , `write_csv` (with a brief introduction to other data import/ export functions from [`readr`](https://readr.tidyverse.org/)).

2.  to take a subset of *columns* in the existing data: `select`

3.  to rename columns: `rename`

4.  to take a subset of *rows* by some simple conditions: `slice_`

5.  to take a subset of *rows* by some more complicated conditions: `filter`

6.  to sort the rows based on the value of one or multiple columns: `arrange`

7.  to perform (4) (5) (6) group by group: `group_by`, `ungroup`

8.  to create new columns in the data: `group_by`, `mutate`, `ungroup`

9.  to summarize the data: `group_by`, `summarise`, `ungroup`


### 0. Load the `tidyverse` Packages

This section loads the packages we need in this lecture.

```{r}
library(tidyverse)
```

### 1. Import and Export the V-Dem Data

This section loads the VDEM dataset and describe its basic information

```{r}
d <- read_csv("_DataPublic_/vdem/1984_2022/vdem_1984_2022_external.csv")
```

### 2. Select economic development indicators

Use the names() function to take a look at all the columns in the dataset

```{r}
names(d)
```

The `str()` provides not only variable names, but also their data types and a few example data points.

```{r}
# Warning: If you have many variables, the output of str() will be lengthy!
str(d)
```

Use the select function to have an overview of the columns of interest

```{r}
d |> select(country_name, country_id, year)
```

Using the `distinct()` function we can effectively identify the distinctive columns
```{r}
d |> select(country_name, country_id, year) |> distinct()
```


```{r}
# Which countries are in this dataset
d |> select(country_name) 

# We see that mexico repeats a lot of times, distinct() will give the names of all the different countries
d |> select(country_name) |> distinct(country_name) 

```

We get all the distinctive years in the dataset

```{r}
d |> select(year) |> distinct() |> print (n = 39)
```

Select both the country identifiers, GDP, and GDP per capita.
It helps us to narrow down the columns in the dataset, does not change the dataset

```{r}
d |> select(country_name, country_id, year, e_gdp, e_gdppc)
```

We create a new dataset in the environment as d_gdp using the assignment operator

```{r}
d_gdp <- d |> 
  select(country_name, country_id, year, e_gdp, e_gdppc)

# If you type the name of the dataset, it will be printed in the R Markdown Document
d_gdp
```

### 3. Rename Columns to Make Names Informative


```{r}

# We are doing nothing to the d_gdp dataset yet, we are just looking at what the output will look like 

d_gdp |>
  rename("GDP" = "e_gdp", "GDP_per_capita" = "e_gdppc",
          "Country" = "country_name", "ID" = "country_id",
          "Year" = "year")

# If we are satisfied with the output, we add d_gdp <- to initiate the change

d_gdp <- d_gdp |>
  rename("GDP" = "e_gdp", "GDP_per_capita" = "e_gdppc",
         "Country" = "country_name", "ID" = "country_id",
         "Year" = "year")

d_gdp
```

### 4. Subset Rows of the Data Using `slice_`

The set of `slice_` functions can be used to select subset of rows using some simple rules

If you would like to get 10 obervations (countries-years) with the maximum `GDP`, use `slice_max`

```{r}
# Want countries-years with highest GDP
# Order by = GDP means that we order countries according to their GDP
# n = 10 means that top 10 countries only

d_gdp |> slice_max(order_by = GDP, n = 10)
```

Similiarily, if you want a subset of countries-years with mimnimal GDP, use `slice_min`:

```{r}
# Get countries-years with the lowest GDP

d_gdp |> slice_min(order_by = GDP, n = 10)
```

If you want to take a random sample of observations in the data, use `slice_sample`. 
Note that you may tell R the exact sample size you want e.g. 10 observations

If you want to get the same random sample everytime, use the set.seed() function and write the same number e.g. 52, to get the same sample. 

This helps makes your code reproducible 

```{r}
set.seed(52)
d_gdp |> slice_sample(n = 10) # Sample 10 observations
```

Or you may define the sample size as a poroportion of the original data size: 10%

```{r}
set.seed(52)
d_gdp |> slice_sample(prop = 0.1)
```


### 5. Subset Rows of the Data Using `filter`

The filter function will take a subset of rows according to our desired condition/criteria

For example, we may take the observations whose `Year` variable ranges from 2000 to 2005.

```{r}
# Want: 2000-2005 data
d_gdp |> filter(Year >= 2000 & Year <= 2005)
```

We may also stack multiple `filter` functions. 
A subset of the data whose `Year` ranges from 2000 to 2005 and another subset where the 
`Country` equals to `"China"

```{r}
# Want: 2000 - 2005 from China
# d_gdp |> filter(Year >= 2000 & Year <= 2005) |> filter(Country == "China")

# If the line becomes to long, we can shift it to a new line after the pipe symbol

d_gdp |> 
  filter(Year >= 2000 & Year <= 2005) |> 
  filter(Country == "China")
```

### 6. Sort the Data based on Values of Rows using `arrange`

We want to arrange the countries according to their GDP per capita, use arrange()

```{r}
# Want: sort the row by GDP per capita
d_gdp |> arrange(GDP_per_capita)

# We got the lowest to highest 
```

If we want the countries-years with larger values of `GDP_per_capita` appear first

1) Add minus before column name
2) Use desc function inside arrange function

```{r}
d_gdp |> arrange(-GDP_per_capita)

d_gdp |> arrange(desc(GDP_per_capita))
```

### 7. Perform (4) (5) (6) group by group: `group_by`, `ungroup`

**Task 1:** Create a dataset that, for each country, include the country-year with the highest GDP.

We want R to do grouping by country, this is where we can find the year with 
the highest gdp 

group_by() causes no change to the dataset, it just separates the dataset into
small groups

whatever function comes after group_by(), it will operate it separately onto
each group

```{r}
# Want: For each country, we want the year with the highest GDP
d_gdp_highest <- d_gdp |>
  group_by(Country) |>
  slice_max(GDP, n = 1)

d_gdp_highest

nrow(d_gdp_highest)
```

Note: As you may recall, there are 181 distinct countries in this dataset. However, the number of observations as output is 341. That means there are duplicates. What may cause duplication? The straightforward way to go (especially for small dataset) is to manually examining the data.

```{r}
d_gdp_highest
```

Upon examination, the cause of duplicates is that some countries or regions have no GDP data. In these cases, there is naturally no way for `slice_` function to find a maximum or a minimum. So it includes all! Hong Kong is an example:

```{r}
d_gdp_highest |> filter(Country == "Hong Kong")
```

To remove these cases, we have two options: (1) we can use `filter` to remove observations whose `GDP` column is `NA`; (2) we can look up `slice_` 's documentation and see if there is a way to remove `NA` cases. I'd go for the latter.

```{r}
# Filter out NA cases with the filter() function
d_gdp_highest <- d_gdp |> 
  group_by(Country) |>
  slice_max(GDP, n = 1) |>
  filter(!is.na(GDP))

# specify arguments for the slice_max() function to remove NA cases
d_gdp_highest <- d_gdp |>
  group_by(Country) |>
  slice_max(GDP, n = 1, na_rm = TRUE)
```

**In-class exercise:** Create a dataset that, for each country, include the country-year with the lowest GDP.

```{r}
d_gdp_lowest <- d_gdp |>
  group_by(Country) |>
  slice_min(GDP, n = 1, na_rm = TRUE)

d_gdp_lowest
```

### 8. Create new columns in the data: `group_by`, `mutate`, `ungroup`

Mutate creates new columns in the dataset 

```{r}
# We created a new column called NEW with each row = 1
d_gdp |> mutate(New = 1)

# Create a new column that is a function of existing column: e.g. GDP
# The New column is exactly same as the GDP column
d_gdp |> mutate(New = GDP)

# The value of the NEW column is the logarithm of GDP or smth else 
d_gdp |> mutate(New = log(GDP))
d_gdp |> mutate(New = log(GDP) + 1)

# Want: New column to be GDP relative to average GDP in the world 1984-2022
d_gdp |> mutate(GDP_over_avg = GDP / mean(GDP))

# Since there is missing data in GDP column, we have to ignore it using na.rm = TRUE
d_gdp |> mutate(GDP_over_avg = GDP / mean(GDP, na.rm = TRUE))

# Want: New column to be GDP relative to average GDP of the country in the world 1984-2022
# Since each country has it own average GDP, then we have to use the group by function
# It will calculate the mean for each group
d_gdp |> 
  group_by(Country) |>
  mutate(GDP_over_avg = GDP / mean(GDP, na.rm = TRUE))
```

**Task:** Add the following economic indicators to the data:

1.  Country-year development level with reference to that of 1984
    = GDP of the year divided by 1984 GDP

2.  Year-on-year economic growth.

```{r}
# Country-year development level with reference to that of 1984.
d_gdp |>
  group_by(Country) |>
  arrange(Year) |>
  mutate(GDP_over_1984 = GDP / first(GDP)) |>
  ungroup() |>
  arrange(Country, Year)

# first()
```

```{r}
# Country-year development level with reference to that of 1984.
d_gdp
```

```{r}
# Year-on-year economic growth.
# We use the lag function
# It helps us find the previous or next values in a vector

d_gdp |>
  group_by(Country) |>
  arrange(Year) |>
  mutate(GDP_yoy_change = GDP - lag(GDP, n = 1)) |>
  ungroup() |>
  arrange(Country, Year)
```

### 9. Summarize the data: `group_by`, `summarise`, `ungroup`

Mutate adds a new column, the number of rows remain the same 

Summarize function changes the number of rows in the dataset 

```{r}
# Want: Average GDP and GDP per capital level of the world
d_gdp |> summarise(gdp_average = mean(GDP, na.rm = TRUE),
                   gdp_per_capita_average = mean(GDP_per_capita, na.rm = TRUE))

# We got the desired summary statistics of the whole dataset 
```

**Task:** Perform a data availability/ integrity check. Then aggregate the data into a new country-level dataset which contains the following indicators:

1.  Average development level from 1984 to 2022.

2.  Magnitude of growth from 1984 to 2022.

```{r}
# Data availability/ integrity check
d_gdp |>
  # Create a column that indicates whether the value is missing
  mutate(GDP_missing = as.numeric(is.na(GDP)), .after = GDP) |>
  group_by(Country) |>
  summarise(N_GDP_missing = sum(GDP_missing))

# ?as.numeric

```

```{r}
# Average development level
d_gdp |>
  group_by(Country) |>
  summarise(GDP_average = mean(GDP, na.rm = TRUE),
            GDPpc_average = mean(GDP_per_capita, na.rm = TRUE))

```

```{r}
# GDP growth and GDP per capita growth: comparing 2019 with 1984
d_gdp |>
  filter(Year >= 1984 & Year <= 2019) |>
  group_by(Country) |>
  arrange(Year) |>
  summarise(GDP_growth_2019_1984 = (last(GDP) - first(GDP)) / first(GDP),
            GDPpc_growth_2019_1984 = (last(GDP_per_capita) - first(GDP_per_capita)) / first(GDP_per_capita)) |>
  ungroup() |>
  arrange(Country)
```

## Final Notes

### Piping

If you break down the result into steps, then you need to have intermediate output

Intermediate output makes the environment messy, so we need to reduce the amount of intermediate output

We use the pipe operator |> to reduce the number of intermediate output, 
it technically pass downs the output to the next command

If you need an intermediate output continuously, then create it 

If you create an intermediate output, but you don't need it subsequently, 
then eliminate it

### Style
Take a look at tidyverse style guide online resource

#### Spaces
Never add space between the function and the argument that follows, function should immediately come with bracket
E.g. filter(Year, XXXX)  vs filter (Year, XXXX)

#### Brackets
Any open bracket should follow a closed bracket 



