---
title: "Note_Data_Wrangling_(2)"
author: "Nisaar Hussain"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Opening Remarks:
We mainly did data wrangling in a single data frame 
Now, we are going to work with multiple data frames  


## Objective

1.  ... to *reshape* a table (long \<-\> wide) with `pivot_longer` and `pivot_wider`
2.  ... to *stack* tables by row or by column with `bind_rows` and `bind_cols` (or, alternatively, `cbind` and `rbind`)
3.  ... to *merge* two tables with `inner_join`, `full_join`, `left_join`, `right_join`, `semi_join`, and `anti_join`


```{r}
library(tidyverse)
```

```{r}
d <- read_csv("_DataPublic_/vdem/1984_2022/vdem_1984_2022_external.csv")

```

### 1. Reshape the V-Dem dataset

- Make this dataset from wide to long using the pivot_long functions
- Transposes the columns into rows    

```{r}
# Want: Each row contain country-year-variable

d |> select(country_name) |> distinct()

# Use the start_with function to select columns that start with e
d_subset <- d |> select(country_name, year, starts_with("e_"))

# We have around 189 columns starting with e and they have their corresponding values 

# We convert all these e columns into rows using pivot_longer functions
d_subset_long <- 
  d_subset |> pivot_longer(cols = starts_with("e_"))
```

- Make the years into columns
- Use pivot_wider function

```{r}
d_subset_wide_year <-
  d_subset_long |> pivot_wider(names_from = year, values_from = value)
```

### 2. Stack multiple subsets of the V-Dem datasets

```{r}
d_VdemIndex <- read_csv("_DataPublic_/vdem/1984_2022/vdem_1984_2022_index.csv")
```

- Bind the 2 datasets by rows

```{r}
d_stack <- bind_cols(d, d_VdemIndex)
?bind_cols
```

```{r}
# Want: Stack two tables by rows?
d_1945_1983 <- read_csv("_DataPublic_/vdem/1945_1983/vdem_1945_1983_external.csv")

```

- Bind the 2 datasets by columns

```{r}
# Bind the 2 columns
d_1945_2022 <- bind_rows(d, d_1945_1983)

# Check if they are successfully stacked
d_1945_2022 |> select(year) |> distinct() |> arrange(year)

```

### 3. Join multiple regional subsets of the V-Dem datasets

```{r}
# INSERT CODE
```

### 4. Validate the GDP data in V-Dem with World Bank data

**Task:** There are many different "versions" of GDP data. I wonder whether the GDP data in the V-Dem dataset is reliable. So I would like to validate it with [data from the World Bank](https://data.worldbank.org/).

#### Download World Bank Data

We will start the adventure by downloading World Bank data.

```{r}
# Install the WDI package that helps fetch data from the World Bank dataset
# See: https://github.com/vincentarelbundock/WDI

# install.packages("WDI")

# Note: Comment out the above "install.packages" command after you are done with installing the package.
```

```{r}
library(WDI)
```

```{r}
# Search for GDP related data
wb_gdpdata_list <- WDIsearch("gdp")

str(wb_gdpdata_list)

# Narrow down to indicators of GDP (I have done some pre-screening)
wb_gdpdata_list_s <- wb_gdpdata_list |> filter(str_detect(indicator, "^NY\\.GDP"))
```

```{r}
# Download GDP-related data
wb_gdpdata <- WDI(
  indicator = c("NY.GDP.MKTP.PP.KD", "NY.GDP.PCAP.PP.KD"), 
  country = "all", 
  start = 1984, end = 2022)
```

```{r}
# Remove the intermediate data we no longer need.
rm(wb_gdpdata_list, wb_gdpdata_list_s)
```

