---
title: "Data Wrangling (3)"
author: "Nisaar Hussain"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

## Outline

-   Reshape (long \<-\> wide) with `pivot_longer` and `pivot_wider`

-   Stack tables by row or by column with `bind_rows` and `bind_cols`

-   Merge two tables with `inner_join`, `full_join`, `left_join`, `right_join`, `semi_join`, and `anti_join`

-   Saving outputs

## Example: The V-Dem Data

\scriptsize

```{r, message=FALSE}
library(tidyverse)
d <- read_csv("_DataPublic_/vdem/1984_2022/vdem_1984_2022_external.csv")
d |> print(n = 3)
```

## Example: The V-Dem Data

-   Focus on the economic indicators: GDP and GDP per capita.

```{r}
d_gdp <- d |> 
  select(country_text_id, year, e_gdp, e_gdppc) |>
  rename("gdp" = "e_gdp", "gdppc" = "e_gdppc")

d_gdp |> print(n = 3)
```

# Reshape a Table

## Wide to Long: `pivot_longer`

-   Convert the columns into rows
-   We want 2 rows for each year, gdp and gdppc

```{r}
d_gdp_long <- d_gdp |>
  pivot_longer(cols = c("gdp", "gdppc"), 
               names_to = "variable", values_to = "value")

d_gdp_long |> print(n = 4)

```

## Long to Wide: `pivot_wider`

Task: Reverse the above `pivot_long` operation.

```{r}
d_gdp_wide_1 <- d_gdp_long |>
  pivot_wider(names_from = "variable", values_from = "value")

d_gdp_wide_1 |> print(n = 4)
```

## Long to Wide: `pivot_wider`

Task: Make `year` the column variable.

- We want to convert the all the years into columns

\scriptsize

```{r}
# Since the dataset needs to get wider, we use pivot_wider function
d_gdp_wide_2 <- d_gdp_long |>
  pivot_wider(names_from = "year", values_from = "value")

d_gdp_wide_2 |> print(n = 2)
```

## Long to Wide: `pivot_wider`

Task: Make `country_text_id` the column variable.
-   Make all the countries as the columns, use pivot_wider

\scriptsize

```{r}
d_gdp_wide_3 <- d_gdp_long |>
  pivot_wider(names_from = "country_text_id", values_from = "value")

d_gdp_wide_3 |> print(n = 2)
```

## Why Do We Reshape Data?

-   **For data cleaning:** Sometime it is much easier to clean the data after reshaping

-   **For data visualization:** Some data visualization functions only take tables shaped in a specific way

-   **For data sharing:** Sometimes you want to export the data for human readers (e.g., data coding/ labeling)


# Stack Tables

## Task

-   We want to merge the data `d_gdp` with some additional datasets that can safely stack together.


##  Load New Data (1)

Make a table with GDP data from two previous time periods (1945-1983 and 1906-1944).

\scriptsize

```{r, message=FALSE}
d_gdp_1945 <- 
  read_csv("_DataPublic_/vdem/1945_1983/vdem_1945_1983_external.csv") |>
  select(country_text_id, year, e_gdp, e_gdppc) |>
  rename("gdp" = "e_gdp", "gdppc" = "e_gdppc")

d_gdp_1906 <- 
  read_csv("_DataPublic_/vdem/1906_1944/vdem_1906_1944_external.csv") |>
  select(country_text_id, year, e_gdp, e_gdppc) |>
  rename("gdp" = "e_gdp", "gdppc" = "e_gdppc")

d_gdp_1945 |> print(n = 2) 
```

## Load New Data (2)

Make two subsets of `d` --- one with education indicators, another with Freedom House indicators.

\scriptsize

```{r}
d_edu <- d |>
  select(e_peaveduc, e_peedgini) |>
  rename("edu_15" = "e_peaveduc", "edu_gini" = "e_peedgini")

d_fh <- d |>
  select(starts_with("e_fh")) |>
  rename("fh_CivilLiberty" = "e_fh_cl", "fh_PoliticalRight" = "e_fh_pr",
         "fh_RuleOfLaw" = "e_fh_rol", "fh_Status" = "e_fh_status")

d_fh |> print(n = 2)
```

## `bind_rows`

-   Stack 2 datasets vertically
-   It will look for columns with same names
-   Only stacks the rows of columns that are included in both datasets 

\scriptsize

```{r}
d_gdp_1945_2022 <- bind_rows(d_gdp, d_gdp_1945)
d_gdp_1945_2022 |> print(n = 3)
unique(d_gdp_1945_2022$year) |> sort()

d_gdp_1945_2022_ue_rows <- bind_rows(
 d_gdp |> select(-gdppc),
 d_gdp_1945 |> select(-gdp)
)
```

## `bind_rows`

\scriptsize

```{r}
d_gdp_1906_2022 <- bind_rows(d_gdp, d_gdp_1945, d_gdp_1906) # can take multiple data frames
d_gdp_1906_2022 |> print(n = 3)
unique(d_gdp_1906_2022$year) |> sort()
```

## `bind_cols`

-   Add more columns to the dataset
-   Add education and freedom house indicators columns to the d_gdp dataset
-   Same syntax as bind_rows
-   Both datasets should have same number of rows or you will get an error message

\scriptsize

```{r}
d_gdp_edu_fh <- bind_cols(d_gdp, d_edu, d_fh) # can take multiple data frames
d_gdp_edu_fh |> print(n = 3)
names(d_gdp_edu_fh)
```

## !! WARNING !! for bind_rows and bind_cols

### These are error-prone operations

-   Do `bind_rows` and `bind_cols` ONLY WHEN you know for sure that there will not be a mismatch!

-   If you have any slightest doubt, don't use them.

# Join Tables

-   X_join 

-   X indicates whose observations we are going to keep

## Tasks

Understand the behavior of different `join_` functions

-   `left_join`: Merge and only keep observations whose identifiers (matching keys) appear in the left-hand-side table.

-   `right_join`: Merge and only keep observations whose identifiers (matching keys) appear in the right-hand-side table.

-   `inner_join`: Merge and only keep observations whose identifiers (matching keys) appear in both tables.

-   `full_join`: Merge and keep observations whose identifiers (matching keys) appear either table.

-   `anti_join`: Filter out observations whose identifiers (matching keys) appear in the right-hand-side table

-   `semi_join`: Filter out observations whose identifiers (matching keys) do not appear in the right-hand-side table

## Task 1: The Case

Join two datasets from the V-Dem data using the above different `join_` functions

-   *GDP* data from **2000-2022**

-   *GDP per capita* data from **1984 to 2010**

## Task 1: Setup

\scriptsize

```{r}
d_gdp_2000_2022 <- d |> filter(year %in% 2000:2022) |>
  select(country_text_id, year, e_gdp) |> rename("gdp" = "e_gdp")

d_gdppc_1984_2010 <- d |> filter(year %in% 1984:2010) |>
  select(country_text_id, year, e_gdppc) |> rename("gdppc" = "e_gdppc")

d_gdp_2000_2022 |> print(n = 2)
d_gdppc_1984_2010 |> print(n = 2)
```

## `left_join`


\scriptsize

```{r}
# We keep the data from 2000-2022 and merge it with 2000 - 2010
# The data from 1984 - 1999 is ignored
d_lj <- d_gdp_2000_2022 |> 
  left_join(d_gdppc_1984_2010, by = c("country_text_id", "year"))

# We get the unique years and sort them
d_lj |> print(n = 2)
unique(d_lj$year) |> sort()
```

## `right_join`

\scriptsize

```{r}
# Keep the data from 1984-2010 and merge it with 2000-2010
# The data from 2011-2022 is ignored since it is right join
d_rj <- d_gdp_2000_2022 |> 
  right_join(d_gdppc_1984_2010, by = c("country_text_id", "year"))

d_rj |> print(n = 2)
unique(d_rj$year) |> sort()
```

## `inner_join`

-   only keep records that appear on both datasets (OVERLAP)
-   Uses "AND"

\scriptsize

```{r}
# Only 2000 - 2010 will appear in the output
d_ij <- d_gdp_2000_2022 |> 
  inner_join(d_gdppc_1984_2010, by = c("country_text_id", "year"))

d_ij |> print(n = 2)
unique(d_ij$year) |> sort()
```

## `full_join`

-   As long as the identifiers appear in 1 dataset, it will be included
-   Uses "OR"

\scriptsize

```{r}
d_fj <- d_gdp_2000_2022 |> 
  full_join(d_gdppc_1984_2010, by = c("country_text_id", "year"))

d_fj |> print(n = 2)
unique(d_fj$year) |> sort()
```

## `semi_join`

-   Check if the identifier appears in the right table
-   If it does, it keeps the records

\scriptsize

```{r}
# Only the data from 2000-2010 will remain
# It find out which columns on the left table are included in the right table
d_sj <- d_gdp_2000_2022 |> 
  semi_join(d_gdppc_1984_2010, by = c("country_text_id", "year"))

d_sj |> print(n = 2)
unique(d_sj$year) |> sort()
```

## `anti_join`

-   If the identifiers appear on right table, then it removes them

\scriptsize

```{r}
# Only show data from 2011 - 2022
# Data from 1984 - 2010 is removed because it is on the RHS
# Data from 2011 - 2022 is left because it does not appear on the RHS
d_aj <- d_gdp_2000_2022 |> 
  anti_join(d_gdppc_1984_2010, by = c("country_text_id", "year"))

d_aj |> print(n = 2)
unique(d_aj$year) |> sort()
```

## Join by Identifiers with Different Variable Names

-   We previously have same column names, but what if the column names are different ?

-   If the identifiers have different names, you have two options: 
    (1) Rename it beforehand 
    (2) specify the `by =` argument differently.

\scriptsize

```{r}
# I make an artificial example whose variable name of a matching 
# identifier is different from d_gdp_2020_2022.
d_gdppc_1984_2010_t <- d_gdppc_1984_2010 |> 
  rename("country_id" = "country_text_id")
```

```{r}
# Option 1: Rename the variables beforehand
d_aj_t <- d_gdp_2000_2022 |> 
  rename("country_id" = "country_text_id") |>
  anti_join(d_gdppc_1984_2010_t, by = c("country_id", "year"))

# Option 2: Specify the "by =" argument with a *named vector*
d_aj_t_2 <- d_gdp_2000_2022 |> 
  anti_join(d_gdppc_1984_2010_t, 
            by = c("country_text_id" = "country_id", 
                   "year" = "year"))

```

## Many-to-One Join: Repeat!

Calculate each country's average 1984-2010 *GDP per capita* and merge it with our annual GDP data from 2000 to 2022.

\scriptsize

```{r}
d_gdppc_1984_2010_avg <- d_gdppc_1984_2010 |> group_by(country_text_id) |>
  summarise(gdppc_1984to2010 = mean(gdppc, na.rm = TRUE))
d_gdppc_1984_2010_avg |> print(n = 2)
```

```{r}
d_lj_ManyToOne <- d_gdp_2000_2022 |> 
  left_join(d_gdppc_1984_2010_avg, by = "country_text_id")
d_lj_ManyToOne |> print(n = 2)
```

## Good Habit

Joining tables is also error-prone.

-   You want to have a clear mind about which variables from which datasets are kept in your final merged data.

-   Failing to do so can cause difficulty with replication.

-   Add suffixes or prefixes indicating data sources 
    (so that you identify the data source directly from the variable)
    (e.g use wb from variable coming from World Bank dataset)

-   Add binary indicators (1/0) indicating from in which dataset is each observation available

## Good Habit: Add Availability Indicators

-   Add binary indicators about data availability in each sources.

\scriptsize

```{r}
# The d_gdp_2000_2022 data are from V-Dem
d_gdp_2000_2022_t <- d_gdp_2000_2022 |> mutate(source_vdem = 1)

# *Pretend* that the d_gdppc_1984_2010 data are from the World Bank
d_gdppc_1984_2010_t <- d_gdppc_1984_2010 |> mutate(source_wb = 1)

d_fj_habit <- d_gdp_2000_2022_t |> 
  full_join(d_gdppc_1984_2010_t, by = c("country_text_id", "year"))

d_fj_habit |> print(n = 3)
```

## Good Habit: Add Availability Indicators

What can you do with these binary indicators? We can know the overlaps of multiple sources.

\scriptsize

```{r}
d_fj_habit |>
  group_by(source_vdem, source_wb) |>
  count()
```

\normalsize

If the overlap looks weird to you, you will know that you need to re-examine the data merging process.

## Good Habit: Add Availability Indicators

Question: Why not just check `NA` in each variables?

Answer: An observation can be missing for two reasons

-   It is in the one of the tables but it does not contain a value.

-   It is not in any of the tables at all.

`join_` make it hard to distinguish between the two scenarios.

## Good Habit: Add prefix or suffix to variable names

-   Previously: Give informative names to variable

-   New advice: Add the source of the variables as part of their names if your final dataset is a combination of many different datasets

## Good Habit: Add prefix or suffix to variable names

\scriptsize

```{r}
d_gdp_2000_2022_rn <- d_gdp_2000_2022 |> 
  rename("vdem_gdp" = "gdp")
  # rename_at(vars(-c("country_text_id", "year")), ~str_c("vdem_", .))

d_gdppc_1984_2010_rn <- d_gdppc_1984_2010 |> 
  rename("wb_gdppc" = "gdppc")
  # rename_at(vars(-c("country_text_id", "year")), ~str_c("wb_", .))

d_fj_habit_2 <- d_gdp_2000_2022_rn |>
  full_join(d_gdppc_1984_2010_rn, by = c("country_text_id", "year"))

d_fj_habit_2 |> print(n = 3)
```

# Save Outputs

## File Formats

You can save your clean data in a variety of formats.

-   `.csv` "comma-separated values," readable by Excel or a text editor

-   `.rds` "R data serialization," readable by R only

-     replace read_csv with write_csv

\scriptsize

```{r}
# Save to a .csv file
write_csv(d_gdp_1945_2022, "Lec_06/2_data_wrangling_3/data/gdp_1945_2002.csv")

# Save to a .rds file
saveRDS(d_gdp_1945_2022, "Lec_06/2_data_wrangling_3/data/gdp_1945_2002.rds")
```

## Reloading saved files

You can re-load saved `.csv` and `.rds` files using `read_csv` and `readRDS` respectively

\footnotesize

```{r}
# Read a .csv file
d_read_1 <- read_csv("Lec_06/2_data_wrangling_3/data/gdp_1945_2002.csv")

# Read a .rds file
d_read_2 <- readRDS("Lec_06/2_data_wrangling_3/data/gdp_1945_2002.rds")
```

## When to use csv/rds file format

-   When to save as `.csv`

    -   Simple data types

    -   Want to manually examine it outside R (e.g., Excel)

    -   Want to share it with non-R users

-   When to save as `.rds`

    -   Complex combination of data types

    -   Simply saving for your future use in R

    -   Large dataset and you want to save space

    -   Text data

If you don't care about looking at the data outside R, `.rds` is a safer option.
