---
title: "Data Visualization 4"
author: "Nisaar Hussain"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

Master data visualization methods for spatial-temporal data with `ggplot`

-   Time series data
    -   One time series
    -   Multiple time series
    
-   Spatial data
    -   Color
    -   Bubble
    -   Size
    
-   Spatial-temporal data

## Load Data

\scriptsize

```{r, message=FALSE}
library(tidyverse)
theme_set(theme_bw()) # Default theme for the whole document

d <- readRDS("Lecture 8/data/wealth_and_health.rds")
d |> print(n = 3)
```

## Recode `region` Indicator

\tiny

```{r}
d <- d |>
  mutate(
    region = case_match(
      region,
      1 ~ "Western Europe", 2 ~ "Northern Europe", 3 ~ "Southern Europe",
      4 ~ "Eastern Europe", 5 ~ "Northern Africa", 6 ~ "Western Africa",
      7 ~ "Middle Africa", 8 ~ "Eastern Africa", 9 ~ "Southern Africa",
      10 ~ "Western Asia", 11 ~ "Central Asia", 12 ~ "Eastern Asia",
      13 ~ "South-Eastern Asia", 14 ~ "Southern Asia", 15 ~ "Oceania",
      16 ~ "North America", 17 ~ "Central America", 18 ~ "South America", 19 ~ "Caribbean",
      .default = NA))
table(d$region)
```

# One Time Series

## Life Expectancy in China: Dots

\tiny

```{r}
# Want: Life Expectancy of China by year using dots
# Create a scatter plot
d |> filter(country_text_id == "CHN") |>
  ggplot(aes(x = year, y = life_expectancy)) + geom_point(alpha = 0.5) +
  labs(x = "Year", y = "Life Expectancy", title = "Life Expectancy in China")
```

## Life Expectancy in China: Line

\tiny

```{r}
# Want: Use a trendline instead of dots
# Use: geom_line vs geom_point
d |> filter(country_text_id == "CHN") |>
  ggplot(aes(x = year, y = life_expectancy)) + geom_line() +
  labs(x = "Year", y = "Life Expectancy", title = "Life Expectancy in China")
```

## Life Expectancy in China: Dots + Trend Line

\tiny

```{r}
# Siutation: Trend line does not consider the years with no data of life expectancy
# Want: both line and points
# Use: geom_point and geom_smotth
d |> filter(country_text_id == "CHN") |>
  ggplot(aes(x = year, y = life_expectancy)) + geom_point(alpha = 0.5) + geom_smooth() +
  labs(x = "Year", y = "Life Expectancy", title = "Life Expectancy in China")
```

# Multiple Time Series (Same Variable)

## Life Expectancy in Eastern Asia (Points)

\tiny

```{r}
# Want: Time series data for multiple East Asian countries
d |> filter(region == "Eastern Asia") |>
  ggplot(aes(x = year, y = life_expectancy, color = country_text_id)) + geom_point(alpha = 0.5) +
  labs(x = "Year", y = "Life Expectancy", title = "Life Expectancy in Eastern Asia", color = "Country/Region")
```

## Life Expectancy in Eastern Asia (Points, Facets)

\tiny

```{r}
# Want: Using facets
d |> filter(region == "Eastern Asia") |>
  ggplot(aes(x = year, y = life_expectancy, color = country_text_id)) + geom_point(alpha = 0.5) +
  facet_wrap(~country_text_id, nrow = 2) +
  labs(x = "Year", y = "Life Expectancy", title = "Life Expectancy in Eastern Asia", color = "Country/Region") + 
  theme(legend.position = "none")
```

## Life Expectancy in Eastern Asia (Points+Trends)

\tiny

```{r}
d |> filter(region == "Eastern Asia") |>
  ggplot(aes(x = year, y = life_expectancy, color = country_text_id)) + geom_point(alpha = 0.2) + geom_smooth() +
  labs(x = "Year", y = "Life Expectancy", title = "Life Expectancy in Eastern Asia", color = "Country/Region")
```

## Life Expectancy in Eastern Asia (Points+Trends, Facets)

\tiny

```{r}
d |> filter(region == "Eastern Asia") |>
  ggplot(aes(x = year, y = life_expectancy, color = country_text_id)) + geom_point(alpha = 0.2) + geom_smooth() +
  facet_wrap(~country_text_id, nrow = 2) +
  labs(x = "Year", y = "Life Expectancy", title = "Life Expectancy in Eastern Asia", color = "Country/Region") + 
  theme(legend.position = "none")
```

## Life Expectancy in Eastern Asia (Annotate Critical Time Point)

\tiny

```{r}
# Situation: Find out which countries have life expectancy over 70 years in year 2000
# Want: Annotated dotted line in 70 and 2000 in each figure
# Use: geom_vline(intercept = 2000) vertical line
# Use: geom_hline(intercept = 70 )  horizontal line 
d |> filter(region == "Eastern Asia") |>
  ggplot(aes(x = year, y = life_expectancy, color = country_text_id)) + geom_point(alpha = 0.2) + geom_smooth() +
  geom_vline(xintercept = 2000, color = "red", linetype = "dashed") + # Add a vertical line at year 2000
  geom_hline(yintercept = 70, color = "red", linetype = "dashed") +  # Add a vertical line at 70 years old
  facet_wrap(~country_text_id, nrow = 2) +
  labs(x = "Year", y = "Life Expectancy", title = "Life Expectancy in Eastern Asia", color = "Country/Region") + 
  theme(legend.position = "none")
```

## Life Expectancy in Eastern Asia (Annotate Critical Time Period)

\tiny

```{r}
# Want: Highlight time period 1984-2020 in yellow 
# Use: annotate, rect = add rectangle in the plot,
# xmin/max = min/max year for the rectangle 
# ymin/max = y-intercept of the rectangle 

d |> filter(region == "Eastern Asia") |>
  ggplot(aes(x = year, y = life_expectancy, color = country_text_id)) + 
  annotate("rect", xmin = 1984, xmax = 2020, ymin = -Inf, ymax = Inf, alpha = 0.3, fill = "yellow", color = "red", linetype = "dashed") +
  geom_point(alpha = 0.2) + geom_smooth() + facet_wrap(~country_text_id, nrow = 2) +
  labs(x = "Year", y = "Life Expectancy", title = "Life Expectancy in Eastern Asia", color = "Country/Region") + 
  theme(legend.position = "none")
```

# Multiple Time Series (Different Variables)

## Evolution of "Health" and "Wealth" in China (Stack --)

\tiny

```{r}
# Situation: R automatically places the order/position of figures in facets
# Want: stack 2 figures of your choice and variables
# E.g. stack health and wealth in china by year figures in 1 plot 
# Step 1: Save each of the figures in an R object and a file
# Step 2: Patchwork allows to put them tgt


library(patchwork) # Package that allow you to "stack" multiple ggplot objects
p_LifeExpectancy <- d |> filter(country_text_id == "CHN") |>
  ggplot(aes(x = year, y = life_expectancy)) + geom_point(alpha = 0.3) + geom_smooth() +
  labs(x = "Year", y = "Life Expectancy", title = "Health in China")
p_gdppc <- d |> filter(country_text_id == "CHN") |>
  ggplot(aes(x = year, y = gdppc)) + geom_point(alpha = 0.3) + geom_smooth() +
  labs(x = "Year", y = "GDP per capita", title = "Wealth in China")
p_LifeExpectancy + p_gdppc # Stack two ggplot objects horizontally with "+"
```

## Evolution of "Health" and "Wealth" in China (Stack \|)

\tiny

```{r}
p_LifeExpectancy / p_gdppc # Stack two ggplot objects vertically with "/"
```

## Evolution of "Health" and "Wealth" in China (Secondary Axis)

**WARNING:** 

This type of figures can be [**confusing**] and [**misleading**]. Unadvised!

\tiny

```{r, fig.height = 3}
coef <- 5
d |> filter(country_text_id == "CHN") |>
  ggplot(aes(x = year)) + 
  geom_line(aes(y = life_expectancy), color = "blue", size = 1) + 
  geom_line(aes(y = gdppc * coef), color = "red", size = 1) +
  scale_y_continuous(name = "Life Expectancy", sec.axis = sec_axis(~./coef, name = "GDP per capita")) +
  theme(axis.title.y.left = element_text(color = "blue"), axis.title.y.right = element_text(color = "red")) +
  labs(x = "Year", title = "GDP")
```
