---
title: "Data Visualization 1"
author: "Nisaar Hussain"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Obejective

-   Data Visualization in R

## Our Task: Extend the "Health and Wealth" Analysis

<https://www.gapminder.org/fw/world-health-chart/>

\vspace{1cm}

> Hans Rosling shows that income and health go hand in hand. People live longer in richer countries. Or the other way around. Countries are richer where people live longer. There are no high income countries with a short life expectancy, and no low income countries with a long life expectancy. Still, there's a huge difference in life expectancy between countries on the same income level, depending on how the money is distributed and how it is used.

# Data Preparation

-   The first step is to prepare your dataset
-   You need to clean it properly
-   Identify columns of interest, rename them, filter out useful year etc.
-   Store the prepared data into a new object for further analysis

## Load the Data

\tiny

```{r, message=FALSE, warning=FALSE}
library(tidyverse)

# Recall, bind_rows functions stacks tables by rows,
# We are stacking the data of different years by rows

d_full <- bind_rows(
  read_csv("_DataPublic_/vdem/1789_1827/vdem_1789_1827_external.csv"),
  read_csv("_DataPublic_/vdem/1867_1905/vdem_1867_1905_external.csv"),
  read_csv("_DataPublic_/vdem/1906_1944/vdem_1906_1944_external.csv"),
  read_csv("_DataPublic_/vdem/1945_1983/vdem_1945_1983_external.csv"),
  read_csv("_DataPublic_/vdem/1984_2022/vdem_1984_2022_external.csv")
)
```

## Select Indicators of Interest

-   Country-year identifiers `country_text_id` `year`

-   Region `e_regiongeo`

-   Wealth: GDP per capita `e_gdppc`

-   Health: Life expectancy `e_pelifeex`

-   Population: `e_mipopula` `e_sb_pop`

-   Extension 1: Infant mortality rate (`e_peinfmor`)

-   Extension 2: Democracy

    -   Binary "democracy" indicator `e_boix_regime`

    -   The Lexical index of Electoral Democracy `e_lexical_index`

    -   Polity V scores (-10 to +10) `e_p_polity`

## Select Indicators of Interest (con'd)

\scriptsize

```{r}
# There are 2 population variables, from different sources
# Recall, we name the variables from the data sources (wb and a research paper)
d <- d_full |>
  select(
    country_text_id, year, 
    e_regiongeo, e_pelifeex, e_gdppc, 
    e_mipopula, e_wb_pop, e_peinfmor,
    e_boix_regime, e_lexical_index, e_p_polity) |>
  rename("region" = "e_regiongeo", 
         "life_expectancy" = "e_pelifeex", 
         "gdppc" = "e_gdppc",
         "population_ClioInfra" = "e_mipopula", 
         "population_WorldBank" = "e_wb_pop",
         "infant_mortality" = "e_peinfmor",
         "democracy_binary" = "e_boix_regime",
         "democracy_lexical" = "e_lexical_index",
         "democracy_polity5" = "e_p_polity") |>
  filter(year >= 1800)

# Saving in R data format ensure replicability
saveRDS(d, "Lecture 7/data/wealth_and_health.rds")
```

## Take a Quick Look at the Data

The summary function to have an overview, check if it looks correct.

-   Look at categorical, quantitative variable 

-   Look at NA, total number of missing values 

-   If there are a lot of missing values, question the usfulness of the data source

\tiny

```{r}
summary(d)
```

## Problem: Multiple Population Data Sources!

We have two population data sources, with different coverage of years.

## Multiple Population Data Sources

Consistency? Check years that are available in both datasets.(Look at codebook)

Both datasets have significant overlaps [1960 - 2000] 


\scriptsize

```{r}
# We compare the population data in the overlapping years
d_pop_overlap <- d |> select(country_text_id, year, starts_with("population_")) |> 
  drop_na() # Remove any row that has at least one missing value
print(d_pop_overlap, n = 3)

# Double checking the overlapping years
unique(d_pop_overlap$year)

# Find the correlation between the two datasets using cor()
cor(d_pop_overlap$population_ClioInfra, d_pop_overlap$population_WorldBank)
```

## Set a Rule to Merge the Two Population Columns

-   Different units: Divide `population_WorldBank` by 1000 
    (so that the unit of population is "in thousands")

-   Different coverage but almost perfect correlation

-   For years that only one dataset has coverage, take the value from the dataset that has available data points.

-   For years that both datasets have coverage, take their `mean`.

That means, effectively, we are taking the `mean` and allow `na.rm = TRUE`. ???

## Merge the Two Population Columns

Below is an implementation of the rule we have just set. 

The output of this step is a new variable called `population`which aggregate data from both sources.

\tiny

```{r}
# STEP 1: "Harmonize" the units
d <- d |> mutate(population_WorldBank = population_WorldBank / 1000)

# STEP 2: Using a non-tidyverse function rowMeans() 
# and create a temporary vector tmp_population, which I remove after use with rm()
# rowMeans() gets the mean of each row, if it is NA, it gets the values from the table without NA

# Creating a temporary vector outside the data frame for inspection
tmp_population <- d |> select(population_ClioInfra, population_WorldBank) |> rowMeans(na.rm = TRUE)
d <- d |> mutate(population = !!(tmp_population), .after = population_WorldBank)

# Remove the vector
rm(tmp_population)

# Remove the columns we no longer need
d <- d |> select(-population_ClioInfra, -population_WorldBank)
```

## Sanity Check

\scriptsize

```{r}
# Take a look at the new population indicator
# There are now far less NA's, so it is better now
summary(d %>% select(-country_text_id, -year, -region))
```

*Always watch out for when you see `NA`, especially when the number is non-trivial!*

## Check Data Availability

\scriptsize

```{r}
# Check for which years we have the required data available
check_data_available <- d |> 
  mutate(Available = (!is.na(life_expectancy) & !is.na(gdppc) & !is.na(population))) 
# Check number of missing values by country-year
table(check_data_available$Available, useNA = "always")

check_data_available |> print(n = 3)
```

## Check Data Availability (con'd)

\scriptsize

```{r}
# More checking, make sure datasets have required information
check_data_available_wide <- check_data_available |>
  select(country_text_id, year, Available) |>
  pivot_wider(names_from = "country_text_id", values_from = "Available", 
              names_prefix = "c_") |>
  arrange(year)

check_data_available_wide |> print(n = 3)
```

## Check Data Availability (con'd)

\scriptsize

```{r}
# Check, for each year, the availability of each column
check_data_available_by_column <- d |>
  group_by(year) |>
  summarise(
    life_expectancy = sum(is.na(life_expectancy)),
    gdppc = sum(is.na(gdppc)),
    population = sum(is.na(population))
  )
  # summarise_at(vars(life_expectancy, gdppc, population), ~sum(!is.na(.))) 
  # above is an alternative way to write the summarise() step

check_data_available_by_column |> print(n = 3)
```

## Save Cleaned Data

**If all looks good, we are ready to create some beautiful data visualization.**

\scriptsize

```{r}
dir.create("Lecture 7/data")
saveRDS(d, "Lecture 7/data/wealth_and_health.rds")
```

# Data Viz Basics

## Simplest Possible Visualization

Plot a scatter plot with ALL the data points.

-   ggplot = plot command

-   aes = specify the variables for x and y axis

-   geom_point = a scatter plot command, plot variables as points

-   labs = labeling for the plot

\tiny

```{r, warning=FALSE, fig.width=9, fig.height=4}
d |> 
  ggplot(aes(x = gdppc, y = life_expectancy)) + 
  geom_point() +
  labs(x = "GDP per capita", y = "Life Expectancy", 
       title = "Wealth and Health in the World (1800-2019)",
       caption = "By Haohan Chen. Data source: V-Dem v.13")
```

## Store Your First Data Visualization

Plots are just like objects created in the environment, we want to save them

To make your nice data visualization stay. You can either (temporarily) save it in your R Environment, or save it as a file in your folder.

\tiny

```{r, warning=FALSE}
# Store in R environment (temporary)
p_all <- d |> 
  ggplot(aes(x = gdppc, y = life_expectancy)) + 
  geom_point() +
  labs(x = "GDP per capita", y = "Life Expectancy", 
       title = "Wealth and Health in the World (1800-2019)",
       caption = "By Nisaar Hussain. Data source: V-Dem v.13")

# Save plot as a .rds file in your folder
saveRDS(p_all, "Lecture 7/Data_Visualization_1/figures/wealth_and_health_all.rds")

# Save plot as a PDF file in your folder, can specify the width and height 
# The quality of figure remains the same in a PDF format despite zooming
ggsave(filename = "Lecture 7/Data_Visualization_1/figures/wealth_and_health_all.pdf", 
       plot = p_all, width = 9, height = 4)

# Save plot as a PNG file, image format
ggsave(filename = "Lecture 7/Data_Visualization_1/figures/wealth_and_health_all.png", 
       plot = p_all)
```

## Themes

- Don't try to change too much of the ggplot default theme settings

- The default design is already good enough 

### Set Themes: `theme_bw`

\tiny

```{r, warning=FALSE, fig.width=9, fig.height=4}
p_all + theme_bw()
```


### Set Themes: `theme_classic`

\tiny

```{r, warning=FALSE, fig.width=9, fig.height=4}
p_all + theme_classic()
```

### Set Themes: `theme_minimal`

\tiny

```{r, warning=FALSE, fig.width=9, fig.height=4}
p_all + theme_minimal()
```

### Other Fancy Themes: The Economist

\tiny

```{r, warning=FALSE, fig.width=9, fig.height=4}
# install.packages("ggthemes") # install the package upon your first use.
# Take a look at the package's website: https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/
library(ggthemes)
p_all + theme_economist()
# p_all + theme_economist_white()
```

### Other Fancy Themes: The WSJ

\tiny

```{r, warning=FALSE, fig.width=9, fig.height=4}
p_all + theme_wsj(base_size = 6)
```
