---
title: "Spatial Data Analysis 1"
author: "Nisaar Hussain"
date: "2023-12-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Required Package

```{r}
library(tidyverse)
library(ggthemes)
theme_set(theme_map())
```

## Case: The Spatial Patterns of Conflict Data

Spatial data analysis on conflicts.

-   [UCDP] Upsala Conflict Data Program: <https://ucdp.uu.se/>

-   [ACLED] Armed Conflict Location & Event Data: <https://acleddata.com/>

## Objective

Explore the simplest way to plot a global map visualizing the country-level frequencies of conflict severity.

## Download UCDP Data

```{r, eval=TRUE}
# Location: https://ucdp.uu.se/downloads/

download.file(
  "https://ucdp.uu.se/downloads/ged/ged221-csv.zip",
  destfile = "Lecture 9/data/ged221-csv.zip")

# Decompress the downloaded ZIP file
unzip("Lecture 9/data/ged221-csv.zip", exdir = "Lecture 9/data")

# Codebook available on the website
```

## Load the Conflict Data

```{r}
d = read_csv("Lecture 9/data/GEDEvent_v22_1.csv")
names(d)
```

## Plot Events on a Map

Plot individual conflict events on a map

Create a dataframe named `d_event_2021` that contains all events in the year of 2021

```{r}
d_event_2021 = d |> filter(year == 2021)
```

## Load the World Map

```{r}
world = map_data("world")
# All the long and lat points are representing the boundaries of the region 

names(world)

# Summary statistics
summary(world)

# Problem: Max. longitude > 180! Remove those
world = world |> filter(long <= 180)
```

```{r}
#Want: filter all the data points for China
world_china = world |> 
  filter(region == "China")

# Use the ggplot to a create map of China based on the data points extracted above
world_china |>
  ggplot() +
  geom_point(aes(x = long, y = lat))

```

Plot the world map

```{r, fig.width=15, fig.height=12}
# Use geom_map function
world |>
  ggplot(aes(x = long, y = lat)) +
  geom_map(
    map = world, # dataset to be used for creating the map
    aes(map_id = region), # using identifier region to carve the boundaries
    color = "white", fill = "gray", linewidth = 0.1
  )

```

**Add map projection**: 

The earth is not a flat plane. 

When we draw maps, we "project" the non-flat plane to a 2D flat plane. 

There are multiple ways through which we can do this projection. For example:

```{r, fig.width=15, fig.height=12}
# use coord_map to create projections of the world map
world |>
  ggplot(aes(x = long, y = lat)) +
  geom_map(
    map = world,
    aes(map_id = region),
    color = "white", fill = "gray", linewidth = 0.1
  ) +
  coord_map() 
```

There are a variety of projections available. 

Choose what you like (for pragmatic and aethetic purpose)

```{r, fig.width=15, fig.height=12}
world |>
  ggplot(aes(x = long, y = lat)) +
  geom_map(
    map = world,
    aes(map_id = region),
    color = "white", fill = "gray", linewidth = 0.1
  ) +
  coord_map("moll")

world |>
  ggplot(aes(x = long, y = lat)) +
  geom_map(
    map = world,
    aes(map_id = region),
    color = "white", fill = "gray", linewidth = 0.1
  ) +
  coord_map("gilbert")

world |>
  ggplot(aes(x = long, y = lat)) +
  geom_map(
    map = world,
    aes(map_id = region),
    color = "white", fill = "gray", linewidth = 0.1
  ) +
  coord_map("orth")


```

## Plot Conflict Events as Points on the map

Plot the events as points on the map.

```{r}
# Visualize civil conflicts in 2021
# Plot only the conflict points, exclude the world map yet
ggplot() +
  geom_point(data = d_event_2021, aes(x = longitude, y = latitude))
```

### Stack the points onto the map

```{r, fig.width=15, fig.height=12}
# First plot the world map as background only
world |>
  ggplot(aes(x = long, y = lat)) +
  geom_map(
    map = world,
    aes(map_id = region),
    color = "white", fill = "gray", linewidth = 0.1
  ) +
  # Below: Stack conflict points onto the map
  geom_point(data = d_event_2021, aes(x = longitude, y = latitude)) +
  coord_map("gilbert")

```

### Make the Plot Prettier

```{r, fig.width=15, fig.height=12}
world |>
  ggplot(aes(x = long, y = lat)) +
  geom_map(
    map = world,
    aes(map_id = region),
    color = "white", fill = "gray", linewidth = 0.1
  ) +
  geom_point(
    data = d_event_2021, 
    aes(x = longitude, y = latitude),
    alpha = 0.3
    ) +
  coord_map()

```

## Plot Conflict Mortality

Use number of deaths as size of points. Calculate total number of deaths

```{r}
# create an indicator specifying the total no. of deaths
d_event_2021 = d_event_2021 |>
  mutate(Deaths = deaths_a + deaths_b + 
           deaths_civilians + deaths_unknown)

summary(d_event_2021$Deaths)

d_event_2021 |>
  ggplot() +
  geom_density(aes(x = Deaths)) +
  theme_bw() +
  scale_x_sqrt() # tackle the skewed distribution as no. of deaths are mainly low

```

```{r, fig.width=15, fig.height=12}
# Want: The size of the points to represent the no. of deaths
# Specify size = Deaths in geom_point
world |>
  ggplot(aes(x = long, y = lat)) +
  geom_map(
    map = world,
    aes(map_id = region),
    color = "white", fill = "gray", size = 0.1
  ) +
  geom_point(
    data = d_event_2021, 
    aes(x = longitude, y = latitude, size = Deaths),
    alpha = 0.3
    ) +
  coord_map() +
  theme(legend.position = "bottom")
```


```{r, fig.width=12, fig.height=8}
# Situation: distribution so skewed that we are not seeing meaningful variations
# Want:Take logarithm of the `Death` variable

world |>
  ggplot(aes(x = long, y = lat)) +
  geom_map(
    map = world,
    aes(map_id = region),
    color = "white", fill = "gray", linewidth = 0.1
  ) +
  geom_point(
    data = d_event_2021, 
    aes(x = longitude, y = latitude, size = log(Deaths+1)),
    alpha = 0.1
    ) +
  coord_map() +
  theme(legend.position = "bottom")

# Q: Why log(Deaths**+1**) ?
```


```{r, fig.width=12, fig.height=8}
# use a more informative color scheme.
world |>
  ggplot(aes(x = long, y = lat)) +
  geom_map(
    map = world,
    aes(map_id = region),
    color = "white", fill = "gray", linewidth = 0.1
  ) +
  geom_point(
    data = d_event_2021, 
    aes(x = longitude, y = latitude, 
        size = log(Deaths+1), color = log(Deaths+1)),
    alpha = 0.1
    ) +
  scale_color_viridis_c(direction = -1, option = "A") +
  coord_map() +
  theme(legend.position = "bottom")
```

