---
title: "Data Visualization 2"
author: "Nisaar Hussain"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

Master data visualization methods with `ggplot`.


## Load Data

\scriptsize

```{r, message=FALSE}
library(tidyverse)
theme_set(theme_bw()) # Default theme for the whole document

d <- readRDS("Lecture 7/data/wealth_and_health.rds")
d |> print(n = 3)
```

## `ggplot` syntax

-   Consider: "Painting" your data on a canvas, layer by layer

-   Steps

    -   Specify the data you want to viaulize

    -   Specify the type(s) of visualization tools

    -   Label the figure (titles of the figure, axes, and legends)

-   Command: Use **`+`** to connect layers (ATTENTION: not `|>` )

Note: [***Always***]{.underline} properly label your figures!

## `ggplot` syntax: example

\scriptsize

```{r, fig.height=3, fig.width=9, warning=FALSE}
d |> # Specify the data frame to plot
  ggplot(aes(x = gdppc, y = life_expectancy)) + # Specify the variables to plot
  geom_point() + # Choose the visualization tool: point
  labs(x = "GDP per capita", y = "Life Expectancy", # set up the axes' titles
       title = "Wealth and Health in the World (1800-2019)", # set up the figure's title
       caption = "By Haohan Chen. Data source: V-Dem v.13") # set up footnote
```

## Variable Types

Visualization tools to use largely depends on 2 variable types

-   "Quantitative", have lots of values (Continuous, Count, Numeric)

    -   GDP per capita

    -   Life expectancy

    -   Population

    -   Infant mortality

-   Categorical, divided into categories 

    -   Binary: Binary "democracy" indicator (1 or 0)

    -   Nominal: Region

    -   Ordinal: Lexical Index of Electoral Democracy

# Quant. X 1

## Viaulize One Quantitative Variable

-   Summary Statistics

    -   Min, max, median, mean, standard deviation, variance,

    -   Quantiles, quintiles

-   Visualization

    -   Histogram

    -   Density Plot

-   Advanced Features

    -   Skewed distribution

    -   Annotate cases

## Summary Statistics

\scriptsize

```{r}
summary(d$gdppc) # Quick summary
length(d$gdppc) # Number of observations
is.na(d$gdppc) |> sum() # Number of missing values
```

## Summary Statistics (con'd)

\scriptsize

```{r}
mean(d$gdppc, na.rm = TRUE) # Mean
var(d$gdppc, na.rm = TRUE) # Variance
sd(d$gdppc, na.rm = TRUE) # Standard deviation
quantile(d$gdppc, na.rm = TRUE) # Quantile
quantile(d$gdppc, seq(0, 1, 0.1), na.rm = TRUE) # Quintile
```

## Histogram

\scriptsize

```{r, fig.width=9, fig.height=3}
d |> 
  ggplot(aes(x = gdppc)) + # Only need to specify x-axis
  geom_histogram() + # Draw a histogram
  labs(x = "GDP per capita", y = "Count", title = "Histogram of GDP per capita")
```

## Histogram (customized)

\scriptsize

```{r, fig.width=10, fig.height=2, warning=FALSE}
# bins = split the data in X bars
# by default bins = 30
d |> ggplot(aes(x = gdppc)) + geom_histogram(bins = 10) +
  labs(x = "GDP per capita", y = "Count", title = "Histogram of GDP per capita")
```

```{r, fig.width=10, fig.height=2, warning=FALSE}
# fill = color of the bars 
# color = border color of the bars
d |> ggplot(aes(x = gdppc)) + geom_histogram(bins = 50, fill = "blue", color = "red") +
  labs(x = "GDP per capita", y = "Count", title = "Histogram of GDP per capita")
```

## Density

-   Used for continuous variables 
-   Analyze the area under the plot

\scriptsize

```{r, fig.width=9, fig.height=3, warning=FALSE}
d |> 
  ggplot(aes(x = gdppc)) +
  geom_density() + # Draw a density plot
  labs(x = "GDP per capita", y = "Density", title = "Density of GDP per capita")
```

## Density (customized)

\scriptsize

```{r, fig.width=9, fig.height=3, warning=FALSE}
d |> 
  ggplot(aes(x = gdppc)) +
  geom_density(fill = "gray", color = "blue") + # Draw a density plot
  labs(x = "GDP per capita", y = "Density", title = "Density of GDP per capita")
```

## Skewed Distribution

-   **Challenge:** Quantitative data may have skewed distributions, which can make the visualization uninformative.

-   **Solution:**

    -   Re-scale the axes

    -   Transform the variable(s)

## Skewed Distribution: Re-scale Axes

\tiny

```{r, fig.width=10, fig.height=2, warning=FALSE, message=FALSE}
# use the scale_x_sqrt()
d |> ggplot(aes(x = gdppc)) + geom_histogram() +
  scale_x_sqrt() + # Scale the x axis to make distribution easy to read
  labs(x = "GDP per capita", y = "Count", title = "Histogram of GDP per capita")
```

```{r, fig.width=10, fig.height=2, warning=FALSE, message=FALSE}
# Take the logarithm of the x-axis
d |> ggplot(aes(x = gdppc)) + geom_histogram() +
  scale_x_log10() + # Scale the x axis to make distribution easy to read
  labs(x = "GDP per capita", y = "Count", title = "Histogram of GDP per capita")
```

## Skewed Distribution: Transform Variables

Alternatively, transform the variable before your visualization

\tiny

```{r, fig.width=10, fig.height=2, warning=FALSE, message=FALSE}
# Change the gdp per cap column to square root before visualization
d |> mutate(gdppc_sqrt = sqrt(gdppc)) |>
  ggplot(aes(x = gdppc_sqrt)) + geom_histogram() + # Proceed to visualization
  labs(x = "sqrt(GDP per capita)", y = "Count", title = "Histogram of GDP per capita")
```

```{r, fig.width=10, fig.height=2, warning=FALSE, message=FALSE}
d |> mutate(gdppc_log = log(gdppc)) |> # Transform the variable before visualization
  ggplot(aes(x = gdppc_log)) + geom_histogram() +
  labs(x = "log(GDP per capita)", y = "Count", title = "Histogram of GDP per capita")
```

## Skewed Distribution: Note

**The method only work when the variable contains NON-NEGATIVE values only.**

\scriptsize

```{r}
# Use the summary function to analyze the negative values, focus on the Min.
summary(d$gdppc)
```

## Annotate Case(s) of Interest

- You want to point out something in the plot, you have to annotate

\tiny

```{r, fig.width=9, fig.height=2, warning=FALSE, message=FALSE}
d |> ggplot(aes(x = gdppc)) + geom_histogram() +
  labs(x = "GDP per capita", y = "Count", title = "Histogram of GDP per capita")
```

## Annotate Case(s) of Interest (Manual Approach)

\tiny

```{r, fig.width=9, fig.height=2, warning=FALSE, message=FALSE}
# Want: Annotate China's gdppc in 2019
# Filter out the required row and store it to new object d_anno
# Create new column "anno" which includes the desired annotation text
d_anno <- d |> filter(country_text_id == "CHN", year == 2019) |> mutate(anno = paste(country_text_id, year), .after = year)
print(d_anno)

# Use the annotate function,hjust = 0 (put the text to the right of the line) 
# geom_vline (vertical line)
d |> ggplot(aes(x = gdppc)) + geom_histogram() + 
  geom_vline(aes(xintercept = 15.4), linetype = "dashed", color = "red") +
  annotate("text", x = 15.4, y = 7500, hjust = 0, label = "China 2019", color = "red") +
  labs(x = "GDP per capita", y = "Count", title = "Histogram of GDP per capita")
# Use the previous figure to get the y = 7500
```

## Annotate Case(s) of Interest (Automated Approach)

\tiny

```{r, fig.width=9, fig.height=2, warning=FALSE, message=FALSE}

# We first use the d_anno dataset for our plot
# Use d_anno dataset in geom_vline and geom_text
# Finally, use the pipe operator to apply in the required dataset 
d |> ggplot(aes(x = gdppc)) + geom_histogram() + 
  geom_vline(data = d_anno, aes(xintercept = gdppc), linetype = "dashed", color = "red") +
  geom_text(data = d_anno, aes(x = gdppc, y = 7500, label = anno), hjust = 0, color = "red") +
  labs(x = "GDP per capita", y = "Count", title = "Histogram of GDP per capita")
```

\normalsize

Why bother with the automated approach ? 

-   It helps when you want to annotate multiple cases.

-   Less prone to error.

## Annotate Case(s) of Interest (Histogram)

\tiny

```{r, fig.width=9, fig.height=2, warning=FALSE, message=FALSE}
# Want: Annotate 3 countries in the histogram plot
d_anno <- d |> filter(country_text_id %in% c("CHN", "JPN", "USA"), year == 2019) |> 
  mutate(anno = paste(country_text_id, year), .after = year)
print(d_anno)

d |> ggplot(aes(x = gdppc)) + geom_histogram() + 
  geom_vline(data = d_anno, aes(xintercept = gdppc), linetype = "dashed", color = "red") +
  geom_text(data = d_anno, aes(x = gdppc, y = 7500, label = anno), hjust = 0, color = "red") +
  labs(x = "GDP per capita", y = "Count", title = "Histogram of GDP per capita")
```

## Annotate Case(s) of Interest (Density)

\tiny

```{r, fig.width=9, fig.height=2, warning=FALSE, message=FALSE}
# Want: Annotate 3 countries in the density plot
d |> ggplot(aes(x = gdppc)) + 
  geom_density() + 
  geom_vline(data = d_anno, aes(xintercept = gdppc), linetype = "dashed", color = "red") +
  geom_text(data = d_anno, aes(x = gdppc, y = 0.15, label = anno), hjust = 0, color = "red") +
  labs(x = "GDP per capita", y = "Count", title = "Histogram of GDP per capita")

# y is changed to 0.15 through observation, not the same as histogram
```

# Quant. X 2

## Visualize Two Quantitative Variables

-   Plots

    -   Scatter plot

    -   Scatter plot with a trend line

-   Advanced

    -   Skewed distribution

    -   Annotate data points

    -   2D "Histogram"

-   Summary Statistics: Correlation

## Scatter Plot

\tiny

```{r, fig.width=9, fig.height=3, warning=FALSE, message=FALSE}
d |> ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_point() +
  labs(x = "GDP per capita", y = "Life Expectancy", title = "Wealth and Health in the World (1800-2019)")
```

## Scatter Plot (customized)

\tiny

```{r, fig.width=9, fig.height=3, warning=FALSE, message=FALSE}
# alpha = transparency of the dots, makes overlapping visible
# shape = can change to triangle etc
d |> ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_point(alpha = 0.3, color = "navy", shape = 3, size = 0.5, stroke = 1) +
  labs(x = "GDP per capita", y = "Life Expectancy", title = "Wealth and Health in the World (1800-2019)")
```

## Scatter Plot with a Trend Line

\tiny

```{r, fig.width=9, fig.height=3, warning=FALSE, message=FALSE}
# Want: Add a trend line to visualize the relationship
# geom_smooth: create a non linear line for the relationship
d |> ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_point(alpha = 0.3, shape = 3, size = 0.5) +
  geom_smooth() +
  labs(x = "GDP per capita", y = "Life Expectancy", title = "Wealth and Health in the World (1800-2019)")
```

## Scatter Plot with a Trend Line (customized aesthetic)

\tiny

```{r, fig.width=9, fig.height=3, warning=FALSE, message=FALSE}
# Customize aesthetics of the smooth line
# se = standard error, shaded areas along the line, specify the color of the line 
d |> ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_point(alpha = 0.3, shape = 3, size = 0.5) +
  geom_smooth(se = FALSE, color = "red") +
  labs(x = "GDP per capita", y = "Life Expectancy", title = "Wealth and Health in the World (1800-2019)")
```

## Scatter Plot with a Trend Line (customized model)

\tiny

```{r, fig.width=9, fig.height=3, warning=FALSE, message=FALSE}
# If you are clear about the relationship
# Want: linear line
# specify [method = "lm"] in geom_smooth

d |> ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_point(alpha = 0.3, shape = 3, size = 0.5) +
  geom_smooth(method = "lm") +
  labs(x = "GDP per capita", y = "Life Expectancy", title = "Wealth and Health in the World (1800-2019)")

# This is a BAD example,because the relationship between the two variables is clearly not linear!
```

\normalsize


## Skewed Distribution: Re-scale Axes

-   1 of the variables is skewed, there will be a skewed distribution. 

\tiny

```{r, fig.width=9, fig.height=3, warning=FALSE, message=FALSE}
# Rescale the x-axis using logarithm
d |> ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_point(alpha = 0.3, color = "navy", shape = 3, size = 0.5, stroke = 1) +
  scale_x_log10() +
  labs(x = "GDP per capita", y = "Life Expectancy", title = "Wealth and Health in the World (1800-2019)")
```

## Skewed Distribution: Transform variables

\tiny

```{r, fig.width=9, fig.height=3, warning=FALSE, message=FALSE}
# Tranform gdppc column to log before visualization
d |> 
  mutate(gdppc_log = log(gdppc)) |>
  ggplot(aes(x = gdppc_log, y = life_expectancy)) +
  geom_point(alpha = 0.3, color = "navy", shape = 3, size = 0.5, stroke = 1) +
  labs(x = "log(GDP per capita)", y = "Life Expectancy", title = "Wealth and Health in the World (1800-2019)")
```

## Annotate Data Points: `geom_text`

\tiny

```{r, fig.width=9, fig.height=3, warning=FALSE, message=FALSE}
# Want: Annotate countries in the dataset d_anno on the scatter plot
# Use geom_text, label = anno (label values under the anno column in d_anno )
# Be careful of the ordering, geom_point is before geom_text
d |> ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_point(alpha = 0.3, shape = 3, size = 0.5) +
  geom_text(data = d_anno, aes(x = gdppc, y = life_expectancy, label = anno), color = "red", size = 2.5) +
  labs(x = "GDP per capita", y = "Life Expectancy", title = "Wealth and Health in the World (1800-2019)")
```

## Annotate Data Points: `geom_label`

-   `geom_text` might not be able to highlight data points enough 
     (especially when the scatter plot has a lot of data points). 

-   `geom_label` can help.

\tiny

```{r, fig.width=9, fig.height=3, warning=FALSE, message=FALSE}
# geom_text makes it look better than geom_text
# Be careful of the ordering, geom_point is before geom_label
d |> ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_point(alpha = 0.3, shape = 3, size = 0.5) +
  geom_label(data = d_anno, aes(x = gdppc, y = life_expectancy, label = anno), color = "red", size = 2.5) +
  labs(x = "GDP per capita", y = "Life Expectancy", title = "Wealth and Health in the World (1800-2019)")
```

## Annotate MANY Data Points: Problem

`geom_label` and `geom_text` will be uninformative annotation if the points you want to annotate cluster in a small area of your figure.

\tiny

```{r, fig.width=12, fig.height=4, warning=FALSE, message=FALSE}
# Want: Annotate China's gdppc from 2010 - 2019
# Problem: labels clustered over each other, makes the graph look bad
d_anno_2 <- d |> filter(year >= 2010, country_text_id == "CHN") |> 
  mutate(anno = paste(country_text_id, year), .after = year)

d |> ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_point(alpha = 0.3, shape = 3, size = 0.5) +
  # geom_point(data = d_anno_2, aes(x = gdppc, y = life_expectancy), color = "red", size = 0.5) +
  geom_label(data = d_anno_2, aes(x = gdppc, y = life_expectancy, label = anno), color = "red", size = 2)
```

## Annotate MANY Data Points: `geom_label_repel`

When you need to annotate a few points that may cluster in a small area of your figure.

-   `geom_label_repel` can adjust the labels' locations to avoid overlaps. 

-    R automatically tackles the overlapping of labels

\tiny

```{r, fig.width=12, fig.height=4, warning=FALSE, message=FALSE}
# A new package
install.packages("ggrepel")
library(ggrepel)

# max.overalaps argument needs to be increased, in case it doesn't work  
# Be careful of the ordering, geom_point is before geom_label
d |> ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_point(alpha = 0.3, shape = 3, size = 0.5) +
  # geom_point(data = d_anno_2, aes(x = gdppc, y = life_expectancy), color = "red", size = 0.5) +
  geom_label_repel(data = d_anno_2, aes(x = gdppc, y = life_expectancy, label = anno), 
                   color = "red", size = 2, max.overlaps = 20) +
  labs(x = "GDP per capita", y = "Life Expectancy", title = "Wealth and Health in the World (1800-2019)")
```

## 2D "Histogram" (Basic)

-   If there are so many data points and scatter plot become uninformative.

-   2D histogram becomes useful. 

-   It cuts the whole space into small grids and calculates how many points fall into
each grid. 

\tiny

```{r, fig.width=9, fig.height=3, warning=FALSE, message=FALSE}
# geom_bin_2d makes the difference 
d |> ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_bin_2d(bins = 80) +
  labs(x = "GDP per capita", y = "Life Expectancy", title = "Wealth and Health in the World (1800-2019)")
```

## 2D "Histogram" (scale transformation)

\tiny

```{r, fig.width=9, fig.height=3, warning=FALSE, message=FALSE}
d |> ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_bin_2d(bins = 80) + scale_x_log10() +
  labs(x = "GDP per capita", y = "Life Expectancy", title = "Wealth and Health in the World (1800-2019)")
```

## 2D "Histogram"(color scheme to better show contrast of values)

Default color scheme (dark to light blue).

`scale_fill_viridis_c` to better visualize the different frequencies


\tiny

```{r, fig.width=9, fig.height=3, warning=FALSE, message=FALSE}
d |> ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_bin_2d(bins = 80) + scale_x_log10() +
  scale_fill_viridis_c() +
  labs(x = "GDP per capita", y = "Life Expectancy", title = "Wealth and Health in the World (1800-2019)")
```

## 2D "Histogram"

\tiny

```{r, fig.width=9, fig.height=3, warning=FALSE, message=FALSE}
d |> ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_hex(bins = 80) + scale_x_log10() +
  labs(x = "GDP per capita", y = "Life Expectancy", title = "Wealth and Health in the World (1800-2019)")
```

## Summary Statistics: Pearson Correlation

Pearson correlation (aka. Pearson's *r*), ranging from -1 to 1

A typical indicator of the correlation between two quantitative variables.

\scriptsize

```{r}
cor(d$life_expectancy, d$gdppc, use = "complete.obs", method = "pearson")
cor.test(d$life_expectancy, d$gdppc, use = "complete.obs", method = "pearson")
```

## Summary Statistics: Pearson Correlation

To use Pearson correlation, the two variables should meet the following assumptions:

-   Continuous

-   Linear relationship

-   No outlier

-   Normally distributed

## Summary Statistics: Spearman Correlation

If the data seem to break at least one of the assumptions for Pearson correlation.

-   You may use Spearman correlation (aka. Spearman's $\rho$).

-   Safer choice to calculate correlation.

\scriptsize

```{r}
cor(d$life_expectancy, d$gdppc, use = "complete.obs", method = "spearman")
cor.test(d$life_expectancy, d$gdppc, use = "complete.obs", method = "spearman")
```

# Quant. X 3

## Visualize 3 Quantitative Variables

-   Correlation Matrix

-   Bubble Plot: 3rd variable visualized with size

-   Make it colorful: 3rd variable visualized with color

## Correlation Matrix (Basic)

-   Put all the plots in a single figure

\tiny

```{r, fig.height=4, message=FALSE, warning=FALSE}
# install.packages("GGally")
library(GGally)
d |>
  select(gdppc, life_expectancy, population) |>
  ggpairs()
```

## Correlation Matrix (Customization)

\tiny

```{r, fig.height=4, message=FALSE, warning=FALSE}
d |> select(gdppc, life_expectancy, population) |>
  ggpairs(
    columnLabels = c("GDP per capita", "Life Expectancy", "Population"), # Label variables
    upper = list(continuous = wrap("cor", method = "spearman", color = "blue")), # Upper triangle: Spearman correlation
    diag = list(continuous = wrap("barDiag", bins = 30, fill = "white", color = "black")), # Diagnal: histogram
    lower = list(continuous = wrap("smooth", alpha = 0.1, color = "gray"))) # Lower triag: + linear trend line
```

## Correlation Matrix

\tiny

```{r, fig.height=4, message=FALSE, warning=FALSE}
d |> select(gdppc, life_expectancy, population) |>
  ggpairs(
    columnLabels = c("GDP per capita", "Life Expectancy", "Population"), # Label variables
    upper = list(continuous = wrap("cor", method = "spearman", color = "blue")), # Upper triangle: Spearman correlation
    diag = list(continuous = wrap("barDiag", bins = 30, fill = "white", color = "black")), # Diagnal: histogram
    lower = list(continuous = wrap("smooth_loess", alpha = 0.1, color = "gray"))) # Lower triag: + linear trend line
```

## Correlation Matrix

\tiny

```{r, fig.height=2, message=FALSE, warning=FALSE}
summary(d$population) # The population variable is skewed. Consider transofmration
d |> select(gdppc, life_expectancy, population) |>
  mutate(population = log(population)) |> # log transformation
  ggpairs(
    columnLabels = c("GDP per capita", "Life Expectancy", "log(Population)"), # Label variables
    upper = list(continuous = wrap("cor", method = "spearman", color = "blue")), # Upper triangle: Spearman correlation
    diag = list(continuous = wrap("barDiag", bins = 30, fill = "white", color = "black")), # Diagnal: histogram
    lower = list(continuous = wrap("smooth_loess", alpha = 0.1, color = "gray"))) # Lower triag: + loess trend line
```

## Play with `ggpairs`

```{r}
install.packages("ggpairs")
library(ggpairs)
```

Check out its documentation: <https://www.rdocumentation.org/packages/GGally/versions/1.5.0/topics/ggpairs>


## Bubble Plot

-   We let the size of the bubble according to a variable

-   E.g. size of the bubble is according to population

\tiny

```{r}
# Want: the size of the bubble = population
# Do: geom_point(aes(size = population))
d |> 
  filter(year == 2019) |>
  ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_point(aes(size = population)) +
  labs(x = "GDP per capita", y = "Life Expectancy", title = "Wealth and Health in the World (2019)")
```

## Bubble Plot (customized)

\tiny

```{r, message=FALSE, warning=FALSE}
# Want: Configure the the labeling of the legend
# use size argument in labs
d |> 
  filter(year == 2019) |>
  ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_point(aes(size = population), alpha = 0.3) +
  labs(x = "GDP per capita", y = "Life Expectancy", size = "Population",
       title = "Wealth and Health in the World (2019)") +
  theme(legend.position = "right") # position of the legend
```

## Bubble Plot (annotated)

\tiny

```{r, message=FALSE, warning=FALSE}
# Want: Annotate countries with the largest population in 2019

# Create a new object where filter top 10 countries by population in 2019 
d_anno_pop <- d |> filter(year == 2019) |> slice_max(order_by = population, n = 10)

# use geom_text to refer to the d_anno_pop
# label argument in geom_text will be the desired annotation
d |> 
  filter(year == 2019) |>
  ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_point(aes(size = population), alpha = 0.3) +
  geom_text(data = d_anno_pop, aes(x = gdppc, y = life_expectancy, label = country_text_id, size = population)) +
  labs(x = "GDP per capita", y = "Life Expectancy", size = "Population",
       title = "Wealth and Health in the World (2019)",
       caption = "Note: World's top 10 most populated countries are annotated.") +
  theme(legend.position = "none") # Remove the legend
```

## Make it Colorful

Strengthen the visual message from the third variable

\tiny

```{r, message=FALSE, warning=FALSE}
# Want: color contrast for population
# add color argument in geom_point
d |> 
  filter(year == 2019) |>
  ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_point(aes(size = population, color = population), alpha = 0.3) +
  labs(x = "GDP per capita", y = "Life Expectancy", 
       size = "Population", color = "Population", fill = "Population",
       title = "Wealth and Health in the World (2019)")
```

## Make it Colorful (customized)

\tiny

```{r, message=FALSE, warning=FALSE}
# scale_color_virdis_c for further customization 
d |> 
  filter(year == 2019) |>
  ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_point(aes(size = population, color = population), alpha = 0.3) +
  scale_color_viridis_c(option = "A", trans = "log", direction = -1) +
  labs(x = "GDP per capita", y = "Life Expectancy", 
       size = "Population", color = "Population", fill = "Population",
       title = "Wealth and Health in the World (2019)")
```

# Quant. X 4+?

## Warning

It is *technically* feasible to plot more than 3 variables in the same figure, but

-   You don't want your figure to be too "busy"

-   "Less is more"

## Add infant mortality variable to the bubble plot

In this case, using color to visualize infant mortality is reasonable.

\tiny

```{r, message=FALSE, warning=FALSE}
# use geom-point function
# use size for population and color for infant mortality 
d |> 
  filter(year == 2019) |>
  ggplot(aes(x = gdppc, y = life_expectancy)) +
  geom_point(aes(size = population, color = infant_mortality), alpha = 0.3) +
  scale_color_viridis_c(option = "A") +
  labs(x = "GDP per capita", y = "Life Expectancy", 
       size = "Population", color = "Infant Mortality",
       title = "Wealth and Health in the World (2019)")
```

## Correlation Matrix (4 variables)

\tiny

```{r, message=FALSE, warning=FALSE}
# Need to use ggpairs
d |> filter(year == 2019) |>select(gdppc, life_expectancy, population, infant_mortality) |>
  mutate(population = log(population)) |>
  ggpairs(
    columnLabels = c("GDP per capita", "Life Expectancy", "log(Population)", "Infant Mortality"), # Label variables
    upper = list(continuous = wrap("cor", method = "spearman", color = "blue")),
    diag = list(continuous = wrap("barDiag", bins = 30, fill = "white", color = "black")), 
    lower = list(continuous = wrap("smooth_loess", alpha = 0.3, color = "gray")))
```
